Total Platform Cyber Protection

### 2016-06-07 paper.tex
Some fixes/edits in Firmware section

% Sai: should this be 'arranged' instead of 'arraigned'?
The layers are arraigned from lowest to highest layer. The lower layer
has more control over the computing system, while the upper layers
have less control. For instance, the Hardware layer has total control
over the physical hardware and memory of the machine, while an
application running in the Application layer can only see a portion of
the machine's memory (because of the virtual memory used by the
Operating System).

@@

Another approach is to change the design of the chip to support
runtime monitoring~\cite{Waksman2010}. In this way, the hardware
itself is monitoring itself, 
% Sai: can we rephrase this^ ? sounds weird
rather than external analysis of
side-channels. However, this has all the same drawbacks of changing
the design of hardware, in addition to not being robust in the face of
a powerful attacker who is able to alter the design.

@@

Firmware is the layer of software that sits and interacts directly
with the hardware. Typically, the firmware is controlled from the
upper layers by device drivers in the Operating System.

Many different hardware contain firmware, for instance the boot BIOS
on a motherboard is a form of firmware, embedded systems often contain
specific firmware, hardware peripherals such as network cards contain
firmware as well. We consider all such areas as firmware.

The key traits of firmware is that, although it is binary code, it is
operating/interacting directly with the hardware, so there are no
OS-style abstraction layers such as virtual memory, well-defined
system calls, or hardware abstraction. Therefore, each firmware is
binary code which is specifically tailored to a specific hardware
platform.

Due to the hardware-specific nature of firmware, \emph{reverse
  engineering} is critical to understanding the firmware. In addition,
reverse engineering is particularly challenging because source code is
frequently unavailable, and vendors' desire to protect their
intellectual property makes them actively obfuscate their
firmware. All this means that, extracting semantically useful
information from firmware is a key first step to performing any
security analysis of the firmware.

@@

Vulnerabilities in the firmware code allow an attacker to execute
arbitrary code with the permissions of the firmware. The class of
vulnerabilities are not unique to firmware, as they are the same class
of vulnerabilities that infect traditional binary software (buffer
overflows, logic flaws, etc.). However, firmware typically has full
control over the entire computing system, therefore vulnerabilities in
firmware are much more severe than those in the application layer. In
addition, if an attacker is able to persist their exploit --- because the
compromise occurs at such a low layer of the computing system --- the
attacker can persist \emph{even after reinstalling the operating
  system or application}.

The attack surface for firmware varies depending on where it is used.
In the most general case, the attacks can come from the upper layers,
specifically the Operating System or Hypervisor layers, as these are
the layers of the computing stack that can directly communicate with the
firmware. However, other firmware can be attacked either directly
through the Hardware layer itself, or through the data being
transmitted through the firmware from the hardware. For instance, a
network card firmware could be attacked through network traffic that
the firmware processes.

@@

Trojan capabilities in the firmware itself can allow a system to be
remotely controlled or accessed. This is similar to a vulnerability,
in that an attacker is able to control the firmware, however the key
difference lies in the fact that Trojan behavior will be intentionally
obfuscated by an attacker, whereas a vulnerability is accidentally
introduced by a benign developer. Therefore, Trojan behavior is 
more difficult to detect, because there is an active adversary who is
intentionally obfuscating the behavior.

Vulnerabilities in the firmware and Trojan capabilities are related
because both require \emph{understanding} the behavior and semantics
of the firmware code. This is why reverse engineering is such a
critical component of securing the threats against the firmware layer:
in order to detect either vulnerabilities or Trojan functionality, an
automated tool must first understand the semantics of the particular
firmware/hardware combination.

Researchers have attempted to identify real-world attacks on firmware
code~\cite{Tsow2006, Cui2010, Duflot2011, Basnight2013, Cui2013,
  Zaddach2013, Zaddach2013a, Maskiewicz2014, Stuttgen2015}. These
attacks are particularly effective, and researchers have shown that
these are able to persist through a reboot and cannot be detected by
the upper layers.

Other approaches have looked at how to prevent the tampering of
firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. This protects from
the threat of tampering or altering the firmware, however it does not
address vulnerable firmware.

Another approach has been to include runtime monitoring into the
firmware, so that compromises can be detected~\cite{Cui2010a,
  Cui2011}. This approach is further complicated by the lack of
abstractions in the firmware environment.

Recent work has looked at the problem of updating the design of the
firmware so that it has security features, such as memory handling and
avoiding information leakage~\cite{Koeberl2014, Kauer2007}. However,
forcing firmware developers to use such features is an open issue.

Very few work has looked at how to reverse engineer and understand the
firmware~\cite{Zaddach2014}. This problem is of particular interest,
as understanding the semantics of the firmware and how it interacts
with hardware is a key challenge.

Recent work has looked at how to detect vulnerabilities in
firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
these has an excellent website\footnote{\url{http://firmware.re}}
that contains the results of their reverse engineering and
vulnerability analysis~\cite{Costin2014}.

Of all the layers, firmware has received the least attention from the
research community. Therefore, we believe that focus on this research
area will yield significant dividends.

The major issue with firmware vulnerabilities is how to decouple
firmware from the hardware, potentially using emulation of the
hardware, in order to aid the analysis of the firmware.

In a similar vein, there is need for automated reverse engineering tools,
including analysis of memory layouts, code modules, even answering the
simple question of what in a firmware binary is code and what is data.

>>>>

% Sai: should this be 'arranged' instead of 'arraigned'?%Faris: I think it should be 'arranged'.
The layers are arraigned from lowest to highest layer. The lower layer
has more control over the computing system, while the upper layers
have less control. For instance, the Hardware layer has total control
over the physical hardware and memory of the machine, while an
application running in the Application layer can only see a portion of
the machine's memory (because of the virtual memory used by the
Operating System).

@@

Another approach is to change the design of the chip to support
runtime monitoring~\cite{Waksman2010}. In this way, the hardware
itself is monitoring itself, 
% Sai: can we rephrase this^ ? sounds weird %Faris : Removing the first 'itself' may solve the problem.
rather than external analysis of
side-channels. However, this has all the same drawbacks of changing
the design of hardware, in addition to not being robust in the face of
a powerful attacker who is able to alter the design.

@@

%Note: I did not remove the previous versions of those paragraphs
%(commented them) so feel free to check the previous versions and 
%revert them if my changes have problems.

%Firmware is the layer of software that sits and interacts directly
%with the hardware. Typically, the firmware is controlled from the
%upper layers by device drivers in the Operating System.

%Faris:
Firmware is the layer of software that sits and interacts directly
with the hardware. Typically, the firmware is controlled by the
device drivers in the upper layers of the Operating System.

%Many different hardware contain firmware, for instance the boot BIOS
%on a motherboard is a form of firmware, embedded systems often contain
%specific firmware, hardware peripherals such as network cards contain
%firmware as well. We consider all such areas as firmware.

%Faris:
Many different hardware contain firmware. For instance: the boot BIOS
on a motherboard, embedded systems that contain
specific firmware, hardware peripherals such as network cards. 
We consider all such areas as firmware.

%The key traits of firmware is that, although it is binary code, it is
%operating/interacting directly with the hardware, so there are no
%%OS-style abstraction layers such as virtual memory, well-defined
%system calls, or hardware abstraction. Therefore, each firmware is
%binary code which is specifically tailored to a specific hardware
%platform.

%Faris:
Although firmware is binary code, it is
operating/interacting directly with the hardware, so there are no
OS-style abstraction layers such as virtual memory, well-defined
system calls, or hardware abstraction which is the key trait of firmware. 
Therefore, each firmware is specifically tailored to a specific hardware platform.

%Due to the hardware-specific nature of firmware, \emph{reverse
%	engineering} is critical to understanding the firmware. In addition,
%reverse engineering is particularly challenging because source code is
%frequently unavailable, and vendors' desire to protect their
%intellectual property makes them actively obfuscate their
%firmware. All this means that, extracting semantically useful
%information from firmware is a key first step to performing any
%security analysis of the firmware.

%Faris:
Due to the hardware-specific nature of firmware, \emph{reverse
  engineering} is critical to understand it. In addition,
reverse engineering is particularly challenging because the source code is
frequently unavailable, and vendors' desire to protect their
intellectual property makes them actively obfuscate their
firmware. This means, extracting semantically useful
information from firmware is the first and key step to perform any
security analysis of the firmware.

@@

%Vulnerabilities in the firmware code allow an attacker to execute
%arbitrary code with the permissions of the firmware. The class of
%vulnerabilities are not unique to firmware, as they are the same class
%of vulnerabilities that infect traditional binary software (buffer
%overflows, logic flaws, etc.). However, firmware typically has full
%control over the entire computing system, therefore vulnerabilities in
%firmware are much more severe than those in the application layer. In
%addition, if an attacker is able to persist their exploit --- because the
%compromise occurs at such a low layer of the computing system --- the
%attacker can persist \emph{even after reinstalling the operating
%	system or application}.

%Faris:
Vulnerabilities in the firmware code allow an attacker to execute
arbitrary code with the permissions of the firmware. The class of
vulnerabilities are not unique to firmware, as they are the same class
of vulnerabilities that infect traditional binary software (buffer
overflows, logic flaws, etc.). However, firmware typically has full
control over the entire computing system, therefore vulnerabilities in
firmware are much more severe than those in the application layer. In
addition, the attacker may remain in the system because the
compromise occurs at such a low layer of the computing system so the
exploit persists \emph{even after reinstalling the operating
system or application}.

%The attack surface for firmware varies depending on where it is used.
%In the most general case, the attacks can come from the upper layers,
%specifically the Operating System or Hypervisor layers, as these are
%the layers of the computing stack that can directly communicate with the
%firmware. However, other firmware can be attacked either directly
%through the Hardware layer itself, or through the data being
%transmitted through the firmware from the hardware. For instance, a
%network card firmware could be attacked through network traffic that
%the firmware processes.

%Faris:
The attack surface for firmware varies depending on where it is used.
In the general case, the attacks can come from the upper layers,
specifically the Operating System or Hypervisor layers, as these are
the layers of the computing stack that can directly communicate with the
firmware. However, other firmware can be attacked either directly
through the Hardware layer, or through the data being
transmitted through the firmware from the hardware. For instance, a
network card firmware could be attacked through network traffic that
the firmware processes.

@@

%Trojan capabilities in the firmware itself can allow a system to be
%remotely controlled or accessed. This is similar to a vulnerability,
%in that an attacker is able to control the firmware, however the key
%difference lies in the fact that Trojan behavior will be intentionally
%obfuscated by an attacker, whereas a vulnerability is accidentally
%introduced by a benign developer. Therefore, Trojan behavior is 
%more difficult to detect, because there is an active adversary who is
%intentionally obfuscating the behavior.

%Faris:
Trojan capabilities in the firmware can allow a system to be
remotely controlled or accessed. This is similar to a vulnerability,
in that an attacker is able to control the firmware, however the key
difference lies in the fact that Trojan behavior can be intentionally
obfuscated by an attacker, whereas a vulnerability can accidentally
introduced by a benign developer. Therefore, Trojan behavior is 
difficult to detect.

%Vulnerabilities in the firmware and Trojan capabilities are related
%because both require \emph{understanding} the behavior and semantics
%of the firmware code. This is why reverse engineering is such a
%critical component of securing the threats against the firmware layer:
%in order to detect either vulnerabilities or Trojan functionality, an
%automated tool must first understand the semantics of the particular
%firmware/hardware combination.

%Faris:
Vulnerabilities in the firmware and Trojan capabilities are related
because both require \emph{understanding} the behavior and semantics
of the firmware code. This is why reverse engineering is such a
critical component of securing the firmware layer against the threats.
In order to detect vulnerabilities or trojan functionality, an
automated tool must understand the semantics of the particular
firmware/hardware combination.

\subsection{Research}
%Faris: Some fixes below. 
Researchers have attempted to identify real-world attacks on firmware
code~\cite{Tsow2006, Cui2010, Duflot2011, Basnight2013, Cui2013,
  Zaddach2013, Zaddach2013a, Maskiewicz2014, Stuttgen2015}. These
attacks are particularly effective, and researchers have shown that
these are able to persist through a reboot and cannot be detected by
the upper layers.

Other approaches are about how to prevent the tampering of
firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. These protect from
the threat of tampering or altering the firmware, however it does not
address vulnerable firmware.

Another approach includes runtime monitoring into the
firmware, so that compromises can be detected~\cite{Cui2010a,
  Cui2011}. This approach is further complicated by the lack of
abstractions in the firmware environment.

Recent work has looked into the problem of updating the design of the
firmware so that it has security features, such as memory handling and
avoiding information leakage~\cite{Koeberl2014, Kauer2007}. However,
forcing firmware developers to use such features is an open issue.

Very few work has looked into how to reverse engineer and understand the
firmware~\cite{Zaddach2014}. This problem is of particular interest,
as understanding the semantics of the firmware and how it interacts
with hardware is a key challenge.

Recent work has looked into how to detect vulnerabilities in
firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
these has an excellent website\footnote{\url{http://firmware.re}}
that contains the results of their reverse engineering and
vulnerability analysis~\cite{Costin2014}.

\subsection{Suggested Focus Areas}
%Faris: Some fixes below.
Of all the layers, firmware has received the least attention from the
research community. Therefore, we believe that focus on this research
area will yield significant dividends.

The major challenge with preventing firmware vulnerabilities is how to decouple
firmware from the hardware, potentially using emulation of the
hardware, in order to aid the analysis of the firmware.

In a similar vein, there is need for automated reverse engineering tools,
including analysis of memory layouts, code modules, even answering the
simple question of what is code in a firmware binary and what is data.
